---
title: Measuring utility-based access to nutrition in Utah County, UT using passive mobile devices and NEMS-S data
author:
  - name: Gregory Macfarlane
    affil: 1
    twitter: greg_macfarlane
    email: gregmacfarlane@byu.edu
    orcid: 0000-0003-3999-7584
    main: true
  - name: Emma Stucki
    affil: 1
    twitter: 
    email: gregmacfarlane@byu.edu
  - name: Alisha Redelfs
    affil: 2
    orcid: 0000-0002-3533-6044
    email: alisha_redelfs@byu.edu
  - name: Lori Spruance
    affil: 2
    orcid: 0000-0003-0254-483X
    email: lori.spruance@byu.edu
affiliation:
  - num: 1
    address: Department of Civil and Construction Engineering, Brigham Young University
  - num: 2
    address: Department of Public Health, Brigham Young University
column_numbers: 3
logoright_name: byu_mono_white.png
logoleft_name: frame.png
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: [packages.bib, poster.bib]
primary_colour: \#002E5D 
secondary_colour: \#FFFFFF
knit: pagedown::chrome_print
---


<!--  Original Abstract
Equitable access to adequate nutrition is a critical concern in many
communities. However, prevailing methods to measure access using distance
thresholds are overly simplistic and lack behavioral support: people do not
always patronize the nearest grocery store, if more affordable or preferable
options exist. Simply considering the proportion of the population living within
a given distance of any grocery store ignores these nuances and could lead to
ineffective policy interventions. In this research, we present an initial
attempt to estimate a grocery store destination choice model for residents in
Utah County, Utah using passive mobile device location data paired with a
recently collected NEMS-S data set representing the local food environment. The
model estimates reveal tradeoffs between value, nutrition quality, and travel
distance segmented by income level. Further, the models allow us to develop a
nutrition accessibility picture of the county that is continuously defined and
sensitive to nutrition qualityin a way that typical models using binary distance
thresholds are not. For example,the proposed measure would allow policy makers
to evaluate the case of a low-income neighborhood whose only accessible grocery
store is a high-priced specialty store, or a neighborhood with multiple suitable
stores just beyond a binary distance threshold. This understanding would in turn
allow more targeted interventions, including grants for store siting or
improvement or increasing transportation accessibility from underserved
neighborhoods to existing stores. Further research is necessary to refine the
model estimates and incorporate travel impedance by multiple transportation
modes.-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'asis',
                      echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")

options(knitr.table.format = "html") 
options(tigris_use_cache = TRUE)

library(modelsummary)
library(tidyverse)
library(targets)
library(here)
library(tigris)
library(ggspatial)

knitr::write_bib(c('posterdown', 'rmarkdown','pagedown', "opentripplanner", "modelsummary", "mlogit"), 'packages.bib')
```

# Introduction

Equitable access to adequate nutrition is a critical concern in many
communities. However, prevailing methods to measure access using distance
thresholds are overly simplistic and lack behavioral support: people do not
always patronize the nearest grocery store, if more affordable or preferable
options exist. Simply considering the proportion of the population living within
a given distance of any grocery store ignores these nuances and could lead to
ineffective policy interventions. 

## Objectives

In this research, we present an initial attempt to estimate a grocery store
destination choice model for residents in Utah County, Utah using passive mobile
device location data paired with a recently collected NEMS-S data set
representing the local food environment. This model allows us to infer the relative
importance of 

1.  Travel distance from home neighborhood to grocery store
2.  Observed behavioral preference for store size and amenities

We then use this model to evaluate access to grocery stores in Utah County.

# Methods

Utility-based access is based in destination choice theory. If a person living in
block group $i$ is choosing to travel to a grocery store, the *utility* of a 
particular store $j$ is


$$
U_{ij} = \beta d_{ij} + \mathbf{\gamma} X_j
$$

Where $d_{ij}$ is the distance or travel cost between $i$ and $j$, $X_j$ is a
vector of attributes of the store (size, brand, etc.), and $\beta, \gamma$ are estimated
coefficients. The total benefit (or access) to all options $J$ is the log-sum of 
the utility of everything in the choice set,

$$
\ln\sum_{j \in J} \exp(U_{ij})
$$
We estimate the coefficients for this measure using three data sets:

  1. A NEMS-S survey [@glanz2007nutrition] conducted on grocery stores in Utah
  County, Utah during Spring 2021 provides attributes of the stores including
  their size and amenities.
  1. Mobile device records obtained from StreetLight Data reveal the home block group of visitors to these
  grocery stores.
  1. Travel paths between block group centroids and grocery stores obtained from
  OpenStreetMap via `opentripplanner` [@opentripplanner]
  
To construct the estimation data set, we sample from the mobile device data to
obtain simulated observed "trips" to grocery stores. Alternative options
are sampled randomly from the other stores --- this results in unbiased but
somewhat inefficient estimates. Models are estimated using the `mlogit` package
for R [@mlogit2020].

# Results

```{r models, include = FALSE}
grocery_models <- read_rds(here("data/grocery_models.rds"))

this_map = c(
    "CAR" = "Drive time",
    "typeConvenience Store" = "Convenience Store",
    "typeOther (Write in a description)" = "Other non-standard",
    "pharmacyTRUE" = "Has pharmacy",
    "ethnicTRUE" = "Ethnic market",
    "merchTRUE" = "Has other merchandise",
    "registers" = "Number of registers",
    "selfchecko" = "Number of self-checkout"
  )
f <- function(x) format(round(x, 1), big.mark=",")
gm <-  list(
  list("raw" = "nobs",  "clean" = "Num.Obs.", "fmt" = f),
  list("raw" = "logLik","clean" = "Log Likelihood", "fmt" = f),
  list("raw" = "rho20", "clean" = "McFadden Rho-Sq", "fmt" = 3)
)
modelsummary(grocery_models)
```


The estimated coefficients are intuitive and highly significant considering
standard confidence intervals. People are less likely to go to stores that are
further away by car, or to stores that are convenience stores or ethnic markets
rather than traditional grocery stores. They are more likely to go to larger
stores (as inferred from the number of cash registers), or stores that include
other kinds of merchandise such as home goods or clothing, all else equal.

```{r model-tab}
modelsummary(
  grocery_models,
  coef_map = this_map, gof_map = gm,
  estimate = "{estimate} ({statistic})", 
  statistic = NULL,
  title = "Store Choice Utility Coeffcients",
  note = "t-statistics in parentheses."
)
```


<br>


```{r logsums-map, messages = FALSE, fig.cap="Access to grocery stores based on the 'Size' model."}
grocery_logsums <- read_rds(here("data/grocery_logsums.rds"))
acsdata <- read_rds(here("data/acsdata.rds"))

bg <- block_groups("UT", "Utah", class = "sf") %>%
  transmute(geoid = GEOID) %>%
  left_join(grocery_logsums, by = c("geoid" = "blockgroup")) %>%
  left_join(acsdata, by = "geoid") %>%
  filter(!is.na(logsum))

ggplot(bg, aes(fill = logsum)) + 
  annotation_map_tile(type = "cartolight", zoom = 10) + 
  geom_sf(lwd = 0, alpha = 0.6)  +
  scale_fill_viridis_b("Logsum Accessibility")
```


When the model coefficients are applied to all block groups in the region, 
the comparative nutrition accessibility picture becomes clear. Large retail
areas show the highest comparative accessibility, while outlying regions 
show comparatively less.

<br>


```{r ls-plot, fig.cap = "Correlation of income and logsum-based access to grocery stores."}
ggplot(bg, aes(x = income, y = logsum, size = households / 1000, weight = households)) + 
  geom_point( alpha = 0.5, show.legend = FALSE) + geom_smooth(show.legend = FALSE) +
  xlab("Median block group household annual income") +
  ylab("Logsum Accessibility") +
  scale_size("Block Group Households (000)") + theme_bw()
```


There is not a clear relationship observable between utility-based access to
grocery stores and traditional sociodemographic characteristics 
revealed by this model. But these preliminary results do not yet consider the
improved mobility that high-income households have. Further research is required.

# Next Steps

The driving time is the only current travel impedance term. This
will be expanded to include other modes of travel, weighted by the mode
availability of the agents.

The NEMS-S contains considerably more attributes than those included in these
preliminary results: produce price and variety, dairy offerings, etc. Investigating 
the effect of these variables on store choice is a major next step.

Neighborhoods with different socioeconomic characteristics are likely to 
show different preferences in their grocery store choice; the model should be
segmented to allow for flexibility and accuracy in these preferences.



# References
